### 2013-04-01 ricardogarfe/final_report/000-community-talks.tex
Ricardo García Fernández Case Studies II First Draft.

\documentclass[11pt]{scrartcl}
\usepackage[parfill]{parskip}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{tabulary}
\usepackage{float}
\usepackage{hyperref}

\graphicspath{{../images/}}

\title{\textbf{Case Studies II - Community Talks}}
\subtitle{Technologies and Contributions}
\author{Ricardo Garc\'ia Fern\'andez}
\date{\today}

\begin{document}

\maketitle

\vfill

\begin{flushright}
    \copyright  2013 Ricardo Garc\'ia Fern\'andez - ricardogarfe [at] gmail [dot] com.

    This work is licensed under a Creative Commons 3.0 Unported License.
    To view a copy of this license visit:

    \url{http://creativecommons.org/licenses/by/3.0/legalcode}.
\end{flushright}

\begin{figure}[h]
    \begin{flushright}	
        \includegraphics{by}
        \label{fig:by}
    \end{flushright}
\end{figure}

\newpage

\tableofcontents

\newpage

\pagenumbering{arabic}

\include{001-apache-community-talk}
\include{002-webkit-community-talk}
\include{003-plan9-community-talk}
\include{004-mozilla-community-talk}
\include{005-gnu-project-community-talk}
\include{006-open-document-community-talk}
\include{007-libre-desktop-laws-community-talk}
\include{008-wikipedia-community-talk}
\include{009-liferay-community-talk}
\include{010-azlinux-migration-commuinty-talk}
\include{011-gnome-community-talk}
\include{012-thunderbird-community-talk}
\include{013-canonical-community-talk}
\include{014-osor-community-talk}
\include{015-kde-community-talk}

\begin{thebibliography}{9}

    \bibitem{thunderbird-talk}
    Thunderbird Community Talk,\\
    Ludovic Hirliman,\\  
    Part 1,\\
    \url{http://blip.tv/mswl/case-studie-ii_quality-assurance-in-thunderbird-and-mozilla-5906218},\\
    Part 2 Questions and Answers,\\
    \url{http://blip.tv/mswl/case-studie-ii_quality-assurance-in-thunderbird-and-mozilla-5906218}

\end{thebibliography}

\end{document}